---
title: "Fitting an antibody landscape"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fitting-an-ablandscape}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ablandscapes)
```
 
This vignette cover the basic approach to fitting an antibody landscape.
 
```{r fitting_ablandscape}
# Get the titers that you want to fit an antibody landcape to, in this case
# a pre-vaccination sample from one of the individuals in the 1997 vaccine 
# cohort
titers <- vaccine1997.titers.pre[4,]

# Get the antigenic coordinates of the antigens that have been titrated in 
# the titers object (in corresponding order)
ag_coords <- h3coords2015[match(names(titers), rownames(h3coords2015)),]

# Fit the antibody landscape
lndscp_fit <- ablandscape.fit(
 titers    = titers,
 coords    = ag_coords,
 bandwidth = 10,
 degree    = 1,
 error.sd  = 1
)
```

The landscape fit object returned from `ablandscape.fit()` can now be used to predict 
the antibody landscape height at different coordinates, using the `predict()` method, 
similarly to other fitting approaches like `lm()`.

In the example below we use it to fit landscape values in a grid and plot a 
corresponding contour plot.

```{r predicting_landscapes}
# Get a grid of titers
lndscp_grid <- list(
  x = seq(from = min(ag_coords[,1])-1, to = max(ag_coords[,1])+1),
  y = seq(from = min(ag_coords[,2])-1, to = max(ag_coords[,2])+1)
)

# Predict the landscape heights along the grid
grid_coords    <- expand.grid(lndscp_grid$x, lndscp_grid$y)
lndscp_heights <- predict(
  lndscp_fit,
  coords     = grid_coords,
  crop2chull = TRUE
)

# Convert the predictions back to a matrix matching x and y
lndscp_grid$z <- matrix(
  lndscp_heights,
  nrow = length(lndscp_grid$x),
  ncol = length(lndscp_grid$y)
)

# There is actually a convenience function that will output the same result
lndscp_grid <- predict_lndscp_grid(lndscp_fit)
```
